Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint

### 2014-11-14 experiments.tex

%
In this section, we give an experimental evaluation of our approach,
highlighting three different aspects.

First, we verify that our NZ-WHO method delivers performance that is
at least on par with the original WHO formulation~\cite{Hariharan12} in terms of
accuracy, while at the same time resulting in large computational
savings (Sect.~\ref{sec:variations}).

Second, we demonstrate that our method can be used for multi-view
object class detection in isolation. It can be applied in a sliding
window fashion and deliver 2D bounding box as well as viewpoint
information. Our method compares on par to state-of-the-art in this
case (Sect.~\ref{sec:exp_iso}).

And third, we show that our method can be used to complement the
detections provided by an existing object class detector, such
as DPM~\cite{Felzenszwalb10} or RCNN~\cite{Girshick14}. In this case,
we show a considerable performance improvement compared to previous
work in the task of joint object class detection and viewpoint
estimation (Sect.~\ref{sec:exp_enrich}).

@@

\begin{table}[!htbp]
    \footnotesize
    \setlength{\tabcolsep}{1pt}
    \centering
\begin{tabular}{|c|c|r|c|r|}
\hline
Methods (AP/MPPE) & before calib.  & time & after calib.~\cite{Aubry14} & time \\
\hline\hline
HOG\cite{Dalal05}     & 72.3 / 65.0           &  31ms  & 60.4 / 50.2                 & 8.7 sec \\ 
WHO\cite{Hariharan12} & 82.1 / 85.4           &  3811ms& 84.4 / 83.0                 & 12.4 sec  \\
WHO-CG                & 81.7 / \textbf{84.9}  &  104ms & 83.7 / 87.3                 & 8.3 sec \\
WHO-CG-Z              & 54.4 / 65.1           &  103ms & \textbf{92.8} / 86.7        & 8.7 sec  \\
% NZC-WHO-Z    & 89.10/\textbf{78.64} &    & 91.15/74.79                  &     \\
NZ-WHO-CG {\em (ours)}  & \textbf{90.0} / 82.8  &   79ms & 90.3 / \textbf{86.8}        & 8.5 sec   \\
\hline
\end{tabular}
\caption{Average Precision(AP) and Mean Precision in Pose Estimation (MPPE) \cite{Lopez-Sastre11} variations of WHO on 3DObject Car dataset\cite{Savarese07}. WHO refers to standard WHO using the method presented in \cite{Hariharan12}, WHO-CG uses iterative Conjugate Gradient method to generate WHO. WHO-CG-Z uses whiten the whole template and zero out textureless region. NZ-WHO-CG is the NZ-WHO which whitens only non-zero cells using iterative Conjugate Gradient method. The time column indicates the time to generate one template. We followed calibration procedure presented in \cite{Aubry14}.}
\label{tab:who_initializations}
\end{table}

@@

This result is encouraging, since our approach reaches a level of
performance that is on par to current state-of-the-art while at the
same time being much faster to train. It takes merely a few minutes to
train, while both ALM~\cite{Xiang12} and DPM-VOC+VP~\cite{Pepik12} are
complex models that optimize non-convex objective functions during
training, which is only made tractable by resorting to delayed
constraint generation in the form of hard negative mining, and can
easily take a day on a single machine. In addition, our approach uses
only rendered images, avoiding the need for real-world training data
with costly bounding box and viewpoint annotations.
%
\begin{table}[!htbp]
    \footnotesize
  \begin{center}
    \begin{tabular}{|c|c|c|c|}
    \hline
     AP/MPPE& Ours & ALM\cite{Xiang12} & VOC-DPM\cite{Pepik12} \\
    \hline\hline
    car & \textbf{99.8}/91.7 &  98.4/93.4 & \textbf{99.8}/\textbf{97.5} \\ 
    bicycle & 93.0/87.4 & 93.0/91.4 & \textbf{98.8}/\textbf{97.5} \\
    \hline
    \end{tabular}
  \end{center}
  \caption{Average Precision(AP) and Mean Precision in Pose Estimation (MPPE) on 3DObject dataset car category. Our method produces high quality 2D-3D matching yet performs on par with state-of-the art detectors.}
  \label{tab:3dobject}
\end{table}

\begin{figure}[h]
    \begin{center}
    \setlength\tabcolsep{0pt}
\begin{tabular}{cc}
  \includegraphics[width=0.45\linewidth]{car_3dobject/7.png} &   
  \includegraphics[width=0.45\linewidth]{car_3dobject/8.png}\\ [-15pt]
%   (a) & (b) \\[0pt]
  \includegraphics[width=0.45\linewidth]{car_3dobject/11.png} &   
  \includegraphics[width=0.45\linewidth]{car_3dobject/12.png}\\ [-5pt]
%  (c) & (d) \\[0pt]
\end{tabular}
\end{center}
\caption{Detection results on 3D Object Classes~\cite{Savarese07}. Original image (left) and detection result overlaid on top with a bounding box and corresponding confidence score(right)}
  \label{fig:3dobject}
\end{figure}

@@

To show such ability, we evaluat our method on the PASCAL3D+
dataset~\cite{Xiang14}. This dataset augments PASCAL 2012 images with
high quality viewpoint annotations thus is ideal to measure pose
estimation. The dataset proposes new metric called Average Viewpoint
Precision (AVP) where it measures the area under viewpoint precision
and detection recall curve. The viewpoints is measured by azimuth
similarity. If the distance between predicted azimuth and ground truth
azimuth is below a certain threshold, the viewpoint is correct.
The baseline methods V-DPM~\cite{Xiang14} and
DPM-VOC+VP~\cite{Pepik12}) reported on the PASCAL3D+ dataset are
variants of DPM~\cite{Felzenszwalb10} where each component of DPM
account for various azimuths. Thus V-DPM and VOC-DPM provide discrete
azimuth only whereas our method provides 3D viewpoint (yaw, pitch,
roll), CAD instance (model index, rendering, depth) and focal length.
%
\begin{figure}[t]
 \begin{center}
    \setlength\tabcolsep{0pt}
    \begin{tabular}{ccc}
    % \includegraphics[width=0.9\linewidth]{tuning} 
   \includegraphics[width=0.33\linewidth]{tuning/1.png} &
   \includegraphics[width=0.33\linewidth]{tuning/2.png} &
   \includegraphics[width=0.33\linewidth]{tuning/3.png} \\[-5pt]
   (a) & (b) & (c)\\
   \includegraphics[width=0.33\linewidth]{tuning/4.png} &
   \includegraphics[width=0.33\linewidth]{tuning/5.png} &
   \includegraphics[width=0.33\linewidth]{tuning/6.png} \\[-5pt]
   (d) & (e) & (f)\\
   \end{tabular}
 \end{center}
 \caption{Effect of fine tuning. (left) original image, (middle) initial detection, (right) continuous fine tuning}
 \label{fig:tuning}
\end{figure}

@@

\begin{figure*}[h]

@@

  \includegraphics[width=0.24\textwidth]{bicycle_cnn/7a.png} &   
  \includegraphics[width=0.24\textwidth]{bicycle_cnn/7b.png} &   
  &\\

@@

\caption{Example enriched bounding boxes. Given R-CNN\cite{Girshick14} detection bounding boxes, our method predicted 2D-3D matching reasonably. On the first column, R-CNN detection bounding boxes overlaid. From the second to the last columns, output of our method given a R-CNN box. Blue boxes are R-CNN output and purple boxes are the tightest bounding box enclosing predicted CAD model}

@@

\begin{table*}[!htbp]

@@

\caption{Average Precision(AP) and Average Viewpoint Precision(AVP) on PASCAL3D+~\cite{Xiang14}. For combined method (* + Ours), we use bounding boxes from * and augment viewpoint using our method. R-CNN and run our method to produce 2D-3D matching. If our method fails to give viewpoint, it predicts the viewpoint to be 0. Note that our method gives high quality metadata such as continuous 3D viewpoint, CAD model (rendering depth) and fine-grained category, whereas base-line methods gives 1D discrete azimuths. Our R-CNN was not an optimized version so the detection AP is lower than the state-of-the-art R-CNN performance}

@@

\begin{figure}[h]
\setlength\tabcolsep{1pt}
\centering
\begin{tabular}{cc}
%  \hline
  \includegraphics[width=0.49\linewidth]{car_cnn4_crop.png} &   
  \includegraphics[width=0.49\linewidth]{car_cnn8_crop.png} \\   
  \includegraphics[width=0.49\linewidth]{car_cnn16_crop.png} &   
  \includegraphics[width=0.49\linewidth]{car_cnn24_crop.png} \\   
%  \hline
\end{tabular}
\caption{Average Precision (AP) and Average Viewpoint Precision (AVP)
  and viewpoint confusion table on PASCAL 12 car validation set using
  R-CNN + Ours (full). From the top left, 4 views, 8 views, 16 views and 24 views. Since our method only augment the object detection, the AP remains the same for all cases. As we increase the viewpoint estimation, the AVP decreases. Note that since we estimate null viewpoint if our method fails to find match, there are more 0 viewpoint predictions (top row of confusion matrices) }
  \label{fig:car_cnn_ap}
\end{figure}

@@

\begin{figure}[h]
\setlength\tabcolsep{5pt}
\begin{tabular}{cc}
  \includegraphics[width=0.45\linewidth]{car_pascal/1.png} &   
  \includegraphics[width=0.45\linewidth]{car_pascal/2.png} \\   
  \includegraphics[width=0.45\linewidth]{car_pascal/3.png} &   
  \includegraphics[width=0.45\linewidth]{car_pascal/4.png} \\   
\end{tabular}
\caption{Stability of our method for overlapping R-CNN detection \cite{Girshick14}. Blue box is the R-CNN detection that is feed into the system as initial detection. Purple box and overlaid rendering is the final output of the system. Note that for various overlapping bounding boxes, our system accurately estimated the reasonable result}
  \label{fig:stability}
\end{figure}

>>>>

%
In this section, we give an experimental evaluation of our approach,
highlighting three different aspects.
%
First, we verify that our NZ-WHO method delivers performance that is
at least on par with the original WHO formulation~\cite{Hariharan12} in terms of
accuracy, while at the same time resulting in large computational
savings (Sect.~\ref{sec:variations}).
%
Second, we demonstrate that our method can be used for multi-view
object class detection in isolation. It can be applied in a sliding
window fashion and deliver 2D bounding box as well as viewpoint
information. Our method compares on par to state-of-the-art in this
case (Sect.~\ref{sec:exp_iso}).
%
And third, we show that our method can be used to complement the
detections provided by an existing object class detector, such
as DPM~\cite{Felzenszwalb10} or RCNN~\cite{Girshick14}. In this case,
we show a considerable performance improvement compared to previous
work in the task of joint object class detection and VP
estimation (Sect.~\ref{sec:exp_enrich}).

@@

\begin{table}[!htbp]
  \footnotesize
  \setlength{\tabcolsep}{1pt}
  \centering
  \begin{tabular}{|c|c|r|c|r|}
    \hline
    Methods (AP/MPPE) & before calib.  & time & after calib.~\cite{Aubry14} & time \\
    \hline\hline
    HOG\cite{Dalal05}     & 72.3 / 65.0           &  31ms  & 60.4 / 50.2                 & 8.7 sec \\ 
    WHO\cite{Hariharan12} & 82.1 / 85.4           &  3811ms& 84.4 / 83.0                 & 12.4 sec  \\
    WHO-CG                & 81.7 / \textbf{84.9}  &  104ms & 83.7 / 87.3                 & 8.3 sec \\
    WHO-CG-Z              & 54.4 / 65.1           &  103ms & \textbf{92.8} / 86.7        & 8.7 sec  \\
    % NZC-WHO-Z    & 89.10/\textbf{78.64} &    & 91.15/74.79                  &     \\
    NZ-WHO-CG {\em (ours)}  & \textbf{90.0} / 82.8  &   79ms & 90.3 / \textbf{86.8}        & 8.5 sec   \\
    \hline
  \end{tabular}
  \caption{Average Precision(AP), Mean Precision in Pose Estimation
    (MPPE)~\cite{Lopez-Sastre11}, and runtime for variations of WHO on
    3D Object Classes cars~\cite{Savarese07}. Please see text for
    details.}
  % WHO refers to standard WHO using the
  % method presented in \cite{Hariharan12}, WHO-CG uses iterative
  % Conjugate Gradient method to generate WHO. WHO-CG-Z uses whiten the
  % whole template and zero out textureless region. NZ-WHO-CG is the
  % NZ-WHO which whitens only non-zero cells using iterative Conjugate
  % Gradient method. The time column indicates the time to generate one
  % template. We followed calibration procedure presented in \cite{Aubry14}.}
  \label{tab:who_initializations}
\end{table}

@@

This result is encouraging, since our approach reaches a level of
performance that is on par to current state-of-the-art while at the
same time being much faster to train. It takes merely a few minutes to
train, while both ALM~\cite{Xiang12} and DPM-VOC+VP~\cite{Pepik12} are
complex models that optimize non-convex objective functions during
training, which is only made tractable by resorting to delayed
constraint generation in the form of hard negative mining, and can
easily take a day on a single machine. In addition, our approach uses
only rendered images, avoiding the need for real-world training data
with costly bounding box and viewpoint annotations.
%
\begin{table}[!htbp]
    \footnotesize
  \begin{center}
    \begin{tabular}{|c|c|c|c|}
    \hline
     AP/MPPE& Ours & ALM\cite{Xiang12} & VOC-DPM\cite{Pepik12} \\
    \hline\hline
    car & \textbf{99.8}/91.7 &  98.4/93.4 & \textbf{99.8}/\textbf{97.5} \\ 
    bicycle & 93.0/87.4 & 93.0/91.4 & \textbf{98.8}/\textbf{97.5} \\
    \hline
    \end{tabular}
  \end{center}
  \caption{Average Precision (AP) and Mean Precision in Pose
    Estimation (MPPE) on 3D Object Classes~\cite{Savarese07} cars.}% Our method produces high quality 2D-3D matching yet performs on par with state-of-the art detectors.}
  \label{tab:3dobject}
\end{table}

\begin{figure}[h]
    \begin{center}
    \setlength\tabcolsep{0pt}
\begin{tabular}{cc}
  \includegraphics[width=0.35\linewidth]{car_3dobject/7.png} &   
  \includegraphics[width=0.35\linewidth]{car_3dobject/8.png}\\ [-15pt]
%   (a) & (b) \\[0pt]
  \includegraphics[width=0.35\linewidth]{car_3dobject/11.png} &   
  \includegraphics[width=0.35\linewidth]{car_3dobject/12.png}\\ [-5pt]
%  (c) & (d) \\[0pt]
\end{tabular}
\end{center}
\caption{Detection results on 3D Object
  Classes~\cite{Savarese07}. Original image (left) and detection
  result overlaid on top (right).}% with a bounding box and corresponding confidence score (right).}
  \label{fig:3dobject}
\end{figure}

@@

To show such ability, we evaluat our method on the PASCAL3D+
dataset~\cite{Xiang14}. This dataset augments PASCAL 2012 images with
high quality viewpoint annotations thus is ideal to measure pose
estimation. The dataset proposes new metric called Average Viewpoint
Precision (AVP) where it measures the area under viewpoint precision
and detection recall curve. The viewpoints is measured by azimuth
similarity. If the distance between predicted azimuth and ground truth
azimuth is below a certain threshold, the viewpoint is correct.
The baseline methods V-DPM~\cite{Xiang14} and
DPM-VOC+VP~\cite{Pepik12}) reported on the PASCAL3D+ dataset are
variants of DPM~\cite{Felzenszwalb10} where each component of DPM
account for various azimuths. Thus V-DPM and VOC-DPM provide discrete
azimuth only whereas our method provides 3D viewpoint (yaw, pitch,
roll), CAD instance (model index, rendering, depth) and focal length.
%
\begin{figure}[t]
 \begin{center}
    \setlength\tabcolsep{0pt}
    \begin{tabular}{ccc}
    % \includegraphics[width=0.9\linewidth]{tuning} 
   \includegraphics[width=0.33\linewidth]{tuning/1.png} &
   \includegraphics[width=0.33\linewidth]{tuning/2.png} &
   \includegraphics[width=0.33\linewidth]{tuning/3.png} \\[-5pt]
%   (a) & (b) & (c)\\
   \includegraphics[width=0.33\linewidth]{tuning/4.png} &
   \includegraphics[width=0.33\linewidth]{tuning/5.png} &
   \includegraphics[width=0.33\linewidth]{tuning/6.png} \\[-5pt]
%   (d) & (e) & (f)\\
   \end{tabular}
 \end{center}
 \caption{Effect of fine tuning. (left) original image, (middle) initial detection, (right) continuous fine tuning}
 \label{fig:tuning}
\end{figure}

@@

\begin{figure*}[t]

@@

%
% sorry no space
%  \includegraphics[width=0.24\textwidth]{bicycle_cnn/7a.png} &   
%  \includegraphics[width=0.24\textwidth]{bicycle_cnn/7b.png} &   
%  &\\

@@

\caption{Example enriched bounding boxes. Given R-CNN\cite{Girshick14}
  detection bounding boxes, our method predicted 2D-3D matching
  reasonably. On the first column, R-CNN detection bounding boxes
  overlaid. From the second to the last columns, output of our method
  given a R-CNN box. Blue boxes are R-CNN output and purple boxes are
  the tightest bounding box enclosing predicted CAD model.}

@@

%
\begin{table*}[t]

@@

\caption{Average Precision(AP) and Average Viewpoint Precision (AVP) on
  PASCAL3D+~\cite{Xiang14}. For combined methods (* + Ours), we use
  bounding boxes from * and augment viewpoint using our method.}
  % R-CNN
  % and run our method to produce 2D-3D matching. If our method fails to
  % give viewpoint, it predicts the viewpoint to be 0. Note that our
  % method gives high quality metadata such as continuous 3D viewpoint,
  % CAD model (rendering depth) and fine-grained category, whereas
  % base-line methods gives 1D discrete azimuths. Our R-CNN was not an
  % optimized version so the detection AP is lower than the
  % state-of-the-art R-CNN performance}

@@

\begin{figure}[h]
\setlength\tabcolsep{1pt}
\centering
\begin{tabular}{cc}
%  \hline
  \includegraphics[width=0.49\linewidth]{car_cnn4_crop.png} &   
  \includegraphics[width=0.49\linewidth]{car_cnn8_crop.png} \\   
  \includegraphics[width=0.49\linewidth]{car_cnn16_crop.png} &   
  \includegraphics[width=0.49\linewidth]{car_cnn24_crop.png} \\   
%  \hline
\end{tabular}
\caption{Average Precision (AP) and Average Viewpoint Precision (AVP)
  and viewpoint confusion table on PASCAL 12 car validation set using
  R-CNN + Ours (full). From the top left, 4 views, 8 views, 16 views
  and 24 views.}
 % Since our method only augment the object detection,
 %  the AP remains the same for all cases. As we increase the viewpoint
 %  estimation, the AVP decreases. Note that since we estimate null
 %  viewpoint if our method fails to find match, there are more 0
 %  viewpoint predictions (top row of confusion matrices) }
  \label{fig:car_cnn_ap}
\end{figure}

@@

\begin{figure}[h]
\setlength\tabcolsep{5pt}
\begin{tabular}{cc}
  \includegraphics[width=0.45\linewidth]{car_pascal/1.png} &   
  \includegraphics[width=0.45\linewidth]{car_pascal/2.png} \\   
  \includegraphics[width=0.45\linewidth]{car_pascal/3.png} &   
  \includegraphics[width=0.45\linewidth]{car_pascal/4.png} \\   
\end{tabular}
\caption{Robustness of our method for against irregular and truncated
  R-CNN detection proposals (cyan).}
  % \cite{Girshick14}. Blue box is the R-CNN detection that is feed into
  % the system as initial detection. Purple box and overlaid rendering
  % is the final output of the system. Note that for various overlapping
  % bounding boxes, our system accurately estimated the reasonable
  % result}
  \label{fig:stability}
\end{figure}
